---
title: "Town Creek RSC 16S, denitrification, nutrients, and more"
author: "Colin G. Finlay"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    df_print: paged
---
# Setup

```{r setup, include=FALSE}
rm(list=ls())
knitr::opts_chunk$set(echo = FALSE, warning=FALSE)
#use to set working directory 
knitr::opts_knit$set(root.dir="~/GitHub/RSC_Microbes_and_Nutrients/analyses/")
```

# Load packages and functions

```{r load packages and functions, include FALSE }
#Set source R tools
source("../bin/DiversityFunctions.R")
source("../bin/MothurTools.R")

#load required packages
require("vegan")
require("tidyverse")
require("reshape2")
require("ecodist")
require("ggpubr")
require("grid")
require("rstatix")
require("NSM3")
require("ggpmisc")
require("funrar")
require("glue")
require("sf")
require("NSM3")
require("gdm")
require("ARTool")
```

```{r Bioconductor install and load}
#if (!require("BiocManager", quietly = TRUE))
   # install.packages("BiocManager")
#BiocManager::install(version = "3.22")

require("BiocManager")
```

```{r Phyloseq install and load}
#source("https://raw.githubusercontent.com/joey711/phyloseq/master/inst/scripts/installer.R",
    #   local = TRUE)

#install_phyloseq(branch = "github")

require("phyloseq")

packageVersion('phyloseq')
```

```{r Standard error and confidence intervals}
#Set Std Err and Conf Int
se <- function(x, ...) {
  sd(x, na.rm = TRUE)/sqrt(length(na.omit(x)))
}
ci <- function(x, ...) {
  1.96 * sd(x, na.rm = TRUE)
}
```

# Load files

```{r read in data files}
# Shared file (OTU) from Mothur Februrary 2025 batch run (original shared filed processed in 'CNH2_AgHC_LD_16S.Rmd')
otu_batch <- read.csv("../data/Mothur/OTU/otu_SharedFile_TC_and_TarPam.csv")
```

# Wrangle

```{r}
# Set row names using mis-formatted column X
rownames(otu_batch) <- otu_batch$X
rownames(otu_batch)

# Remove column X now that row names are set
otu_batch <- otu_batch[,-1]
```

# QC OTUs per sample

```{r Visualize Town Creek and Tar Pam samples}
# 1. Check dimensions
dim(otu_batch) # 139, 164,672

# 2. Prepare data for plotting
d_otu_4plot <- as.data.frame(otu_batch)
d_otu_4plot$colsum <- rowSums(d_otu_4plot) 
d_otu_4plot$group <- rownames(d_otu_4plot)

# 3. Plot and save OTUs per sample figure
p_TC_OTUs <- ggplot(d_otu_4plot, aes(x = reorder(group, -colsum), y = colsum)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  # Add the count labels
  geom_text(aes(label = colsum), hjust = -0.2, size = 3.5) + 
  coord_flip() +
  theme_minimal() +
  # Expand the y-axis (which is horizontal now) slightly to make room for the labels
  scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
  labs(x = "", y = "", title = "")

p_TC_OTUs

#ggsave("../figures/MothurCheck_Feb26_batch/TC_and_TarPam_OTUs.pdf", plot=p_TC_OTUs, device="pdf", scale=1, width= 200, height= 500, units = "cm", dpi=600, limitsize=F, bg="white")
```
```{r min and max OTU counts}
# 1. Calculate row sums (excluding the first column if it's the ID)
# Assuming SampleID is column 1 and the rest are counts:
total_counts <- rowSums(otu_batch) 

# 2. Find the index of the max and min values
max_index <- which.max(total_counts)
min_index <- which.min(total_counts)

# 3. Extract the Sample IDs using the index
max_sample_id <- rownames(otu_batch)[max_index]
min_sample_id <- rownames(otu_batch)[min_index]

# Print results
print(paste("Largest Sample ID:", max_sample_id, "| Total:", total_counts[max_index])) # GSF1058_TarPam_02 | Total: 400458
print(paste("Smallest Sample ID:", min_sample_id, "| Total:", total_counts[min_index])) # GSF1058_TC_48 | Total: 70
```

```{r Drop all TarPam samples}
# 1. Identify rows that contain "_TarPam_"
# grepl returns TRUE for matches, FALSE for non-matches
is_tarpam <- grepl("_TarPam_", rownames(otu_batch))

# 2. Subset to keep only the rows that do NOT have the pattern
otu_noTarPam <- otu_batch[!is_tarpam, ]

# 3. Sanity Check: Verify the counts
print(paste("Original rows:", nrow(otu_batch)))
print(paste("Rows removed:", sum(is_tarpam)))
print(paste("Remaining rows:", nrow(otu_noTarPam)))
```

```{r Visualize Town Creek samples}
# 1. Check dimensions
dim(otu_noTarPam) # 112, 164,672

# 2. Prepare data for plotting
d2_otu_4plot <- as.data.frame(otu_noTarPam)
d2_otu_4plot$colsum <- rowSums(d2_otu_4plot) 
d2_otu_4plot$group <- rownames(d2_otu_4plot)

# 3. Plot and save OTUs per sample figure
p_TC_OTUs_2 <- ggplot(d2_otu_4plot, aes(x = reorder(group, -colsum), y = colsum)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  # Add the count labels
  geom_text(aes(label = colsum), hjust = -0.2, size = 3.5) + 
  coord_flip() +
  theme_minimal() +
  # Expand the y-axis (which is horizontal now) slightly to make room for the labels
  scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
  labs(x = "", y = "", title = "")

p_TC_OTUs_2

#ggsave("../figures/MothurCheck_Feb26_batch/TC_OTUs.pdf", plot=p_TC_OTUs_2, device="pdf", scale=1, width= 200, height= 400, units = "cm", dpi=600, limitsize=F, bg="white")
```

Lowest 4 OTU counts are all water (GSF1058_TC_48, GSF1058_TC_50, GSF1058_TC_51, GSF1058_TC_49), and ≤ 425 OTUs. From there, 5th lowest is 7319 OTUs. Plan to drop all water samples before rarefying.

## Rarefaction

```{r rarefaction}
#Rarefy
#Set min sample number
min.N <- min(rowSums(otu4))
min.N # 9 OTUs

# new df wich drops the sample with only 2 otu hits:
otu5 <- otu4[rownames(otu4) != "AGLD004bc093", ]

# new min sample number
min.N_3 <- min(rowSums(otu5))
min.N_3 # 10 OTUs

#graph of read counts
d2<-as.data.frame(otu5)
d2$colsum <- colSums(t(d2))
d2$group <- rownames(d2)
p3<-ggplot(d2, aes(x=group,y=colsum))+geom_bar(stat="identity")+coord_flip()
p3
```

```{r Gemini rarefaction}
# --- Start here with your 'otu2' data frame ---

# 3. Find the minimum sample size (library depth) for rarefaction
# We calculate the total counts for each sample (row)
sample_sums <- rowSums(otu3)

# We find the smallest sample sum to use as our depth
min_depth <- min(sample_sums)

# It's good practice to print this depth
print(paste("Rarefying all samples to a depth of:", min_depth, "reads"))

# 4. Set a random seed for reproducibility
# Rarefaction involves random subsampling. Setting a seed ensures
# that you get the exact same rarefied data every time you run the code.
set.seed(42) 

# 5. Perform rarefaction
# The rrarefy() function takes the community data (otu2) and
# the subsampling depth (min_depth)
otu_rarefied <- rrarefy(otu3, sample = min_depth)

# 6. (OPTIONAL) Check the new row sums to confirm
print("New row sums after rarefaction:")
print(rowSums(otu_rarefied))

# 'otu_rarefied' is your new data frame with rarefied counts
```

Relative abundance using funrar:
- use the package "funrar", and function make_relative() to do relative abundance calculations
  - yields same results as our for loops
  - has publication: https://doi.org/10.1111/ddi.12629 
- install.packages("funrar")
- require("funrar")
- relative_abundance <- make_relative(as.matrix(Numerical))
```{r relative abundance}
otu_rel <- make_relative(as.matrix(otu_rarefied))
```

```{r read in meta data}
#Read in design file
meta <- read.csv("../data/Design_HC_LD.csv")
```

```{r wrangle meta data}
# remove underscores form meta sampleIDs and make row names
rownames(meta) <- gsub(pattern = "_", replacement = "", x = meta$SampleID)

# remove barcodes from otu_rel row names
# This selects and keeps only the first 7 characters of each row name
rownames(otu_rel) <- substr(rownames(otu_rel), start = 1, stop = 7)

# Get the row names from 'otu_rel' (the data frame with fewer rows)
matching_samples <- rownames(otu_rel)

# 1. Filter and reorder 'meta' to match 'otu_rel'
meta_filtered <- meta[matching_samples, ]

# 2. Now cbind the two data frames, which have identical row names and order
meta_otu <- cbind(meta_filtered, otu_rel)

# (Optional) Check the dimensions to confirm
dim(meta_otu) 
# This should show 49 rows
```

```{r set factors}
meta_otu$Site <- as.factor(meta_otu$Site)
meta_otu$Field_or_Buffer <- as.factor(meta_otu$Field_or_Buffer)
```

# PERMANOVA

```{r PERMANOVA}
# PERMANOVA:
metaotu.ad <- adonis2(meta_otu[,-c(1:5)] ~ Site * Field_or_Buffer, method = "bray", data = meta_otu, perm=10000, set.seed=42, by = "terms")
metaotu.ad

metaotu.ad.2 <- adonis2(meta_otu[,-c(1:5)] ~ Site * Field_or_Buffer * Treatment_salt, method = "bray", data = meta_otu, perm=10000, set.seed=42, by = "terms")
metaotu.ad.2 # The interaction between field/buffer and Treatment_salt makes this model unusable

metaotu.ad.3 <- adonis2(meta_otu[,-c(1:5)] ~ Treatment_salt, method = "bray", data = meta_otu, perm=10000, set.seed=42, by = "terms")
metaotu.ad.3 # Treatment_salt overlaps with field/buffer categories too much, not very informative

# For treatment_salt (Hyde) and maybe Field_Location (if not redundant), analyze within Hyde (and maybe within Pitt)
```

# Denitrification and soil parameters

```{r denit wrangling}
# remove NO_ACET
HC_denit_acet <- HC_denit %>%
  dplyr::filter(Acetylene == "ACET", .preserve = TRUE)

LD_denit_acet <- LD_denit %>%
  dplyr::filter(Acetylene == "ACET", .preserve = TRUE)

# for HC, remove extra samples 1 through 4
# Create a vector of the SampleIDs to exclude
values_to_remove <- c("AG_HC_001_ACET", "AG_HC_002_ACET", "AG_HC_003_ACET", "AG_HC_004_ACET")

HC_denit_acet_filtered <- HC_denit_acet[!(HC_denit_acet$Sample_ID %in% values_to_remove), ]

# first, need to simplify so column number and names match
HC_denit_simple <- HC_denit_acet_filtered[,c(1,6)]
LD_denit_simple <- LD_denit_acet[, c(1,8)]

# rbind HC and LD
comb_denit_simple <- rbind(HC_denit_simple, LD_denit_simple)

# row names to match PCoA
# get rid of underscores
rownames(comb_denit_simple) <- gsub(pattern = "_", replacement = "", x = comb_denit_simple$Sample_ID)

# This selects and keeps only the first 7 characters of each row name
rownames(comb_denit_simple) <- substr(rownames(comb_denit_simple), start = 1, stop = 7)
```

```{r soil wrangling}
#Wet_dry_mass
# add dry_proportion (dry/wet)
Wet_dry_mass$dry_proportion <- Wet_dry_mass$dry_soil_g / Wet_dry_mass$wet_soil_g

# add percent soil moisture
Wet_dry_mass$PercentMoisture <- ((Wet_dry_mass$wet_soil_g - Wet_dry_mass$dry_soil_g) / Wet_dry_mass$wet_soil_g) * 100

#N_inorganic
# add dry_proportion column
# replace NAs in N inorganic with zeros
N_inorganic[, c("NH4_uM", "NH4_mg.L", "NO3_NO2_uM", "NO3_NO2_mg.L")] <- lapply(
  N_inorganic[, c("NH4_uM", "NH4_mg.L", "NO3_NO2_uM", "NO3_NO2_mg.L")],
  function(x) ifelse(is.na(x), 0, x))

N_inorganic$dry_proportion <- Wet_dry_mass$dry_proportion

# Calculate dry soil input for KCl extraction (dry soil input = wet g * dry_proportion)
N_inorganic$dry_soil_g <- N_inorganic$KCl_field_moist_g_soil * N_inorganic$dry_proportion

# divide mg/L by dry g for final unit to use in simplified data frame
N_inorganic$NH4_mg.L.gDM <- N_inorganic$NH4_mg.L / N_inorganic$dry_soil_g

N_inorganic$NO3_NO2_mg.L.gDM <- N_inorganic$NO3_NO2_mg.L / N_inorganic$dry_soil_g

#pH
  # already good

#C_N_Sal_EC
# C:N ratio
C_N_Sal_EC$CN <- C_N_Sal_EC$C_perc / C_N_Sal_EC$N_perc
```

```{r simplify, rbind, row names}
# first, need to simplify so column number and names match
Soil_moisture_simp <- Wet_dry_mass[,c(1,5)]
N_inorganic_simp <- N_inorganic[, c(1,9,10)]

# row names to match PCoA
# get rid of underscores
rownames(Soil_moisture_simp) <- gsub(pattern = "_", replacement = "", x = Soil_moisture_simp$SampleID)
rownames(N_inorganic_simp) <- gsub(pattern = "_", replacement = "", x = N_inorganic_simp$Sample)
rownames(pH) <- gsub(pattern = "_", replacement = "", x = pH$SampleID)
rownames(C_N_Sal_EC) <- gsub(pattern = "_", replacement = "", x = C_N_Sal_EC$Sample)

# This selects and keeps only the first 7 characters of each row name
rownames(Soil_moisture_simp) <- substr(rownames(Soil_moisture_simp), start = 1, stop = 7)
rownames(N_inorganic_simp) <- substr(rownames(N_inorganic_simp), start = 1, stop = 7)
rownames(pH) <- substr(rownames(pH), start = 1, stop = 7)
rownames(C_N_Sal_EC) <- substr(rownames(C_N_Sal_EC), start = 1, stop = 7)
```

```{r combine all and filter for envfit}
# combine denit and all soil data frames (cbind)
denit_soil_all <- cbind(comb_denit_simple, Soil_moisture_simp, N_inorganic_simp, pH, C_N_Sal_EC)

# remove sample ID columns
denit_soil_all <- denit_soil_all[,-c(1, 3, 5, 8, 10)]

# create new data frame that only contains 16S samples. This will be the envfit data frame
denit_soil_filtered <- denit_soil_all[rownames(denit_soil_all) %in% rownames(otu_rel), ]
```

```{r soil stat assumptions}
# denit_soil_all with meta data:
denit_soil_all_meta <- cbind(meta, denit_soil_all)

denit_soil_all_meta$log_denit <- log10(denit_soil_all_meta$Denitrification_rate)
denit_soil_all_meta$log_EC <- log10(denit_soil_all_meta$EC_uS.cm)
denit_soil_all_meta$log_moisture <- log10(denit_soil_all_meta$PercentMoisture)

# Normality:
denit_soil_all_meta %>%
  shapiro_test(Denitrification_rate)
  # Not normally distributed

denit_soil_all_meta %>%
  shapiro_test(log_denit)
  # Normally distributed!

denit_soil_all_meta %>%
  shapiro_test(EC_uS.cm)
  # Not normally distributed

denit_soil_all_meta %>%
  shapiro_test(log_EC)
  # Not normally distributed

denit_soil_all_meta %>%
  shapiro_test(PercentMoisture)
  # Not normally distributed

denit_soil_all_meta %>%
  shapiro_test(log_moisture)
  # Not normally distributed

denit_soil_all_meta %>%
  shapiro_test(N_perc)
  # Not normally distributed

denit_soil_all_meta %>%
  shapiro_test(C_perc)
  # Not normally distributed

denit_soil_all_meta %>%
  shapiro_test(CN)
  # Not normally distributed

denit_soil_all_meta %>%
  shapiro_test(NH4_mg.L.gDM)
  # Not normally distributed

denit_soil_all_meta %>%
  shapiro_test(NO3_NO2_mg.L.gDM)
  # Not normally distributed

denit_soil_all_meta %>%
  shapiro_test(mean_pH)
  # Not normally distributed

  # QQ plots:
ggqqplot(denit_soil_all_meta, "log_denit") # May be acceptable

ggqqplot(denit_soil_all_meta, "log_EC") # Bad, but borderline acceptable

ggqqplot(denit_soil_all_meta, "log_moisture") # no parametric possible

ggqqplot(denit_soil_all_meta, "CN")
  # Shapiro failed but QQ looks acceptable for parametric stats

```

```{r ART ANOVA}
denit_soil_all_meta$Site <- as.factor(denit_soil_all_meta$Site)
denit_soil_all_meta$Field_or_Buffer <- as.factor(denit_soil_all_meta$Field_or_Buffer)

m.art = art(Denitrification_rate ~ Site*Field_or_Buffer, data = denit_soil_all_meta)
anova(m.art)
```

# Soil boxplots

## Denit, Moisture, EC 3-panel

```{r plotting}
# add combined site and field or buffer
denit_soil_all_meta$Site_field_buff <- paste(denit_soil_all_meta$Site, denit_soil_all_meta$Field_or_Buffer, sep = "_")

denit_soil_all_meta$Site_field_buff <- as.factor(denit_soil_all_meta$Site_field_buff)

# Kruskal-Wallis test on Treatment, and multiple comparisons:
denit_kruskal<- compare_means(Denitrification_rate~Site_field_buff, denit_soil_all_meta, method = "kruskal.test")

denit_multipleComparisons<- compare_means(Denitrification_rate~Site_field_buff, denit_soil_all_meta, method = "wilcox.test", p.adjust.method = "fdr")

# Adjustments for plotting:
denit_multipleComparisons <- add_significance(denit_multipleComparisons, p.col = "p.adj", output.col = "p.adj.signif", cutpoints = c(0, 1e-04, 0.001, 0.01, 0.05, Inf), symbols = c("****", "***", "**", "*", "ns"))

# Turn into data frame:
denit_MultComp.DF <- as.data.frame(denit_multipleComparisons)

# SDCFlig
SDCFlig_denit <- pSDCFlig(x = denit_soil_all_meta$Denitrification_rate, g = denit_soil_all_meta$Site_field_buff, method = "Monte Carlo", n.mc = 10000)
SDCFlig_denit

# Add on SDCFlig p-values:
denit_MultComp.DF$SDCF_p <- c(SDCFlig_denit$p.val)

# Symbols for plotting:
denit_MultComp.DF <- add_significance(denit_MultComp.DF, p.col = "SDCF_p", output.col = "p.SDCF.signif", cutpoints = c(0, 1e-04, 0.001, 0.01, 0.05, Inf), symbols = c("****", "***", "**", "*", "ns"))

# Letters for plotting
denit_letters <- data.frame(
  Site_field_buff = c("HydeCo_buffer", "HydeCo_field", "PittCo_buffer", "PittCo_field"),
  Letters = c("a", "a", "b", "c")
)


# Export the data frames to a .csv file
#write.csv(N2O_MultComp.DF, file = "../figures/pub/tables/Kruskal_Treatment/multiple_comparisons/N2O_MultComp.csv", row.names = T)

# Weird 'hide.ns' behavior patch:
#N2O_MultComp.DF$p[4] <- 1.0

# Plotting:
plot.denit <- ggplot(denit_soil_all_meta, aes(x=Site_field_buff, y=Denitrification_rate)) +
  geom_boxplot(outlier.shape = NA) +
  geom_point(size = 2, position = position_jitterdodge()) +
  theme_bw() +
  #Remove plot grid lines
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
  #Set axis title and text properties, tick marks, and labels
  theme(text=element_text(size=20),axis.title=element_text(size=20,face="bold"),
          axis.text=element_text(size=20),  
          axis.title.y=element_text(margin=margin(r=10)),
          axis.text.x = element_text(angle = 45, hjust = 1), panel.border = element_rect(colour = "black",size=1), strip.text = element_text(size = 20), legend.position="right", legend.title = element_text(size=20), legend.text=element_text(size=20)) + 
          theme(axis.ticks.length=unit(0.3,"cm")) + 
          labs(x = "", y = expression(paste("ng N"[2],"O g"^{-1}, " h"^{-1})), title="") +
  #scale_x_discrete(labels=c("Dry", "Interim", "Wet"))+
  theme(rect=element_rect(fill="transparent"))+
  theme(plot.background = element_rect(color=NA))+
  #stat_pvalue_manual(data = denit_MultComp.DF, comparisons = list(c("HydeCo_buffer", "HydeCo_field"), c("HydeCo_buffer", "PittCo_buffer"), c("HydeCo_buffer", "PittCo_field"), c("HydeCo_field", "PittCo_buffer"), c("HydeCo_field", "PittCo_field"), c("PittCo_buffer", "PittCo_field")), label = "p.adj.signif", size = 10, vjust = 1, hide.ns = "p", y.position = max(denit_soil_all_meta$Denitrification_rate), step.increase = 0.1)
  geom_text(data = denit_letters, aes(x = Site_field_buff, y = max(denit_soil_all_meta$Denitrification_rate) + 5000, label = Letters), size = 8)

 
plot.denit

#ggsave("../figures/denit_boxplot.png", plot=plot.denit, device="png", path=NULL, scale=1, width=9, height=5, dpi=600, limitsize=TRUE, bg="white")
```

```{r denit site facet plot}
# Rename factor levels
denit_soil_all_meta$Site <- factor(denit_soil_all_meta$Site,
                                   levels = c("HydeCo", "PittCo"),
                                   labels = c("Hyde County", "Pitt County"))
# Letters for plotting
denit_letters.1a <- data.frame(
  Site = c("Hyde County", "Hyde County", "Pitt County", "Pitt County"),
  Field_or_Buffer = c("buffer", "field", "buffer", "field"),
  Letters = c("a", "a", "b", "c")
)


# Plotting:
plot.denit.1a <- ggplot(denit_soil_all_meta, aes(x=Field_or_Buffer, y=Denitrification_rate)) +
  facet_wrap(vars(Site))+
  geom_boxplot(outlier.shape = NA) +
  geom_point(size = 2, position = position_jitterdodge()) +
  theme_bw() +
  #Remove plot grid lines
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
  #Set axis title and text properties, tick marks, and labels
  theme(text=element_text(size=20),axis.title=element_text(size=20,face="bold"),
          axis.text=element_text(size=20),  
          axis.title.y=element_text(margin=margin(r=10)),
          axis.text.x = element_text(angle = 45, hjust = 1), panel.border = element_rect(colour = "black",size=1), strip.text = element_text(size = 20), legend.position="right", legend.title = element_text(size=20), legend.text=element_text(size=20)) + 
          theme(axis.ticks.length=unit(0.3,"cm")) + 
          labs(x = "", y = expression(paste("ng N"[2],"O g"^{-1}, " h"^{-1})), title="") +
  theme(rect=element_rect(fill="transparent"))+
  theme(plot.background = element_rect(color=NA))+
  theme(axis.text.x = element_blank())+
  theme(axis.ticks.x = element_blank())+
  coord_cartesian(ylim = c(-4, 26000))+
  geom_text(data = denit_letters.1a, aes(x = Field_or_Buffer, y = max(denit_soil_all_meta$Denitrification_rate) + 4000, label = Letters), size = 8)

 
plot.denit.1a
```

```{r moisture stats and plot}
# Kruskal-Wallis test on Treatment, and multiple comparisons:
moisture_kruskal<- compare_means(PercentMoisture~Site_field_buff, denit_soil_all_meta, method = "kruskal.test")

# Plotting:
plot.moist.b <- ggplot(denit_soil_all_meta, aes(x=Field_or_Buffer, y=PercentMoisture)) +
  facet_wrap(vars(Site))+
  geom_boxplot(outlier.shape = NA) +
  geom_point(size = 2, position = position_jitterdodge()) +
  theme_bw() +
  #Remove plot grid lines
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
  #Set axis title and text properties, tick marks, and labels
  theme(text=element_text(size=20),axis.title=element_text(size=20,face="bold"),
          axis.text=element_text(size=20),  
          axis.title.y=element_text(margin=margin(r=10)),
          axis.text.x = element_text(angle = 45, hjust = 1), panel.border = element_rect(colour = "black",size=1), strip.text = element_text(size = 20), legend.position="right", legend.title = element_text(size=20), legend.text=element_text(size=20)) + 
          theme(axis.ticks.length=unit(0.3,"cm")) + 
          labs(x = "", y = expression(paste("% soil moisture")), title="") +
  theme(rect=element_rect(fill="transparent"))+
  theme(axis.text.x = element_blank())+
  theme(axis.ticks.x = element_blank())+
  theme(plot.background = element_rect(color=NA))

 
plot.moist.b
```

```{r EC stats and plot}
# Kruskal-Wallis test on Treatment, and multiple comparisons:
EC_kruskal<- compare_means(EC_uS.cm~Site_field_buff, denit_soil_all_meta, method = "kruskal.test")

EC_multipleComparisons<- compare_means(EC_uS.cm~Site_field_buff, denit_soil_all_meta, method = "wilcox.test", p.adjust.method = "fdr")

# Adjustments for plotting:
EC_multipleComparisons <- add_significance(EC_multipleComparisons, p.col = "p.adj", output.col = "p.adj.signif", cutpoints = c(0, 1e-04, 0.001, 0.01, 0.05, Inf), symbols = c("****", "***", "**", "*", "ns"))

# Turn into data frame:
EC_MultComp.DF <- as.data.frame(EC_multipleComparisons)

# Letters for plotting
EC_letters <- data.frame(
  Site_field_buff = c("HydeCo_buffer", "HydeCo_field", "PittCo_buffer", "PittCo_field"),
  Letters = c("a", "a", "b", "c")
)

# SDCFlig
SDCFlig_EC <- pSDCFlig(x = denit_soil_all_meta$EC_uS.cm, g = denit_soil_all_meta$Site_field_buff, method = "Monte Carlo", n.mc = 10000)
SDCFlig_EC

# Add on SDCFlig p-values:
EC_MultComp.DF$SDCF_p <- c(SDCFlig_EC$p.val)

# Symbols for plotting:
EC_MultComp.DF <- add_significance(EC_MultComp.DF, p.col = "SDCF_p", output.col = "p.SDCF.signif", cutpoints = c(0, 1e-04, 0.001, 0.01, 0.05, Inf), symbols = c("****", "***", "**", "*", "ns"))

# Letters for plotting
EC_letters.1a <- data.frame(
  Site = c("Hyde County", "Hyde County", "Pitt County", "Pitt County"),
  Field_or_Buffer = c("buffer", "field", "buffer", "field"),
  Letters = c("ab", "a", "b", "c")
)

# Plotting:
plot.EC.c <- ggplot(denit_soil_all_meta, aes(x=Field_or_Buffer, y=EC_uS.cm)) +
  facet_wrap(vars(Site))+
  geom_boxplot(outlier.shape = NA) +
  geom_point(size = 2, position = position_jitterdodge()) +
  theme_bw() +
  #Remove plot grid lines
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
  #Set axis title and text properties, tick marks, and labels
  theme(text=element_text(size=20),axis.title=element_text(size=20,face="bold"),
          axis.text=element_text(size=20),  
          axis.title.y=element_text(margin=margin(r=10)),
          axis.text.x = element_text(angle = 45, hjust = 1), panel.border = element_rect(colour = "black",size=1), strip.text = element_text(size = 20), legend.position="right", legend.title = element_text(size=20), legend.text=element_text(size=20)) + 
          theme(axis.ticks.length=unit(0.3,"cm")) + 
          labs(x = "", y = expression(paste("EC (µS cm"^{-1}, ")")), title="") +
  theme(rect=element_rect(fill="transparent"))+
  theme(plot.background = element_rect(color=NA))+
  coord_cartesian(ylim = c(0, 3200))+
  geom_text(data = EC_letters.1a, aes(x = Field_or_Buffer, y = max(denit_soil_all_meta$EC_uS.cm) + 500, label = Letters), size = 8)

 
plot.EC.c

```

```{r}
denit_moist_EC_3panel <- ggarrange(plot.denit.1a, plot.moist.b, plot.EC.c, ncol = 1, nrow = 3, align = "hv",labels = c("(a)","(b)","(c)"), legend = "none")

denit_moist_EC_3panel

#ggsave("../figures/denit_moist_EC_3panel.png", plot=denit_moist_EC_3panel, device="png", path=NULL, scale=1, width=10, height=12, dpi=600, limitsize=TRUE, bg="white")
```


# PCoA

```{r Calculate distance matrix and PCoA components without field, end samples only}
# Bray-Curtis distance matrix:
otu.rel.dist<-vegdist(otu_rel, method="bray")

# Principal Coordinates Analysis
otu.rel.dist.pcoa <- cmdscale(otu.rel.dist, k=2, eig=TRUE, add=TRUE)
# Classical (Metric) Multidimensional Scaling; returns PCoA coordinates
# eig=TRUE returns eigenvalues; k = # of dimensions to calculate

explainvar1a <- round(otu.rel.dist.pcoa$eig[1] / sum(otu.rel.dist.pcoa$eig), 3) * 100
explainvar2a <- round(otu.rel.dist.pcoa$eig[2] / sum(otu.rel.dist.pcoa$eig), 3) * 100
sum.eiga <- sum(explainvar1a, explainvar2a)

explainvar1a # 15.8%
explainvar2a # 6%
sum.eiga # 21.8%
```

```{r PCoA data frames}
pcoa.groups <- paste(meta_otu$Site, meta_otu$Field_or_Buffer, sep = "_")
pcoa.points <- data.frame(otu.rel.dist.pcoa$points, group = pcoa.groups)

# Calculate Centroids (mean and SE)
pcoa.L.centroids <- reshape2::melt(pcoa.points, id="group", measure.vars = c("X1", "X2"))
pcoa.centroids <- acast(pcoa.L.centroids, variable ~ group, mean)
pcoa.centroids.se <- acast(pcoa.L.centroids, variable ~ group, se)
pcoa.centroids.sd <- acast(pcoa.L.centroids, variable ~ group, sd)

# Combine
pcoa.cent.dataframe <- cbind(t(pcoa.centroids), t(pcoa.centroids.se))
colnames(pcoa.cent.dataframe) <- c("V1", "V2", "V1e", "V2e")

# Adding columns with the factors. H: hydrologic history
pcoa_Site <- c("HydeCo", "HydeCo", "PittCo", "PittCo")
pcoa_field_or_buff <- c("buffer", "field", "buffer", "field")

# create and add to "trts" data frame
pcoa.cent.dataframe.trts <- as.data.frame(pcoa.cent.dataframe) 
pcoa.cent.dataframe.trts$Site <- as.factor(pcoa_Site)
pcoa.cent.dataframe.trts$Field_or_Buffer <- as.factor(pcoa_field_or_buff)
pcoa.cent.dataframe.trts$centroid_or_sample <- "centroid"

# Add in points to data frame:
pcoa.points_ForPlot <- pcoa.points
colnames(pcoa.points_ForPlot) <- c("V1", "V2", "group")

# Add labels to match ...cent.dataframe.trts
pcoa.points_ForPlot$Site <- meta_otu$Site
pcoa.points_ForPlot$Field_or_Buffer <- meta_otu$Field_or_Buffer

#remove "group" column
pcoa.points_ForPlot <- pcoa.points_ForPlot[, -3]
#add V1e and V2e as zeros:
pcoa.points_ForPlot$V1e <- numeric(49)
pcoa.points_ForPlot$V2e <- numeric(49)
# Add column to distinguish sample points from centroids, and later use to set alpha in plot
pcoa.points_ForPlot$centroid_or_sample <- "sample"

#Merge:
centroids_points_merged <- rbind(pcoa.cent.dataframe.trts, pcoa.points_ForPlot)

# Add combined site and field or buffer column:
centroids_points_merged$Site_field_buff <- paste(centroids_points_merged$Site, centroids_points_merged$Field_or_Buffer, sep = "_")

centroids_points_merged$Site_field_buff <- as.factor(centroids_points_merged$Site_field_buff)
```

```{r envfit}
# pcoa points data frame for envfit
pcoa_DF <- as.data.frame(otu.rel.dist.pcoa$points)

#Run envfit with new Soil data (SSF = Soil Start & Finish):
fit <- envfit(pcoa_DF, denit_soil_filtered, permutations = 10000, set.seed(42))
fit

A_fit <-as.list(fit$vectors)
vec <- as.data.frame(fit$vectors$arrows*sqrt(fit$vectors$r)*1)
p_fit <- as.data.frame(A_fit$pvals)
vec$pvals <- A_fit$pvals
vec <- subset(vec, pvals<=0.05)

vec$labels <- c("%moisture", "EC")
```

```{r PCoA plot site and field or buffer}
#Plot using ggplot2
df <- as.data.frame(centroids_points_merged)

plot_pcoa <- ggplot(df, aes(x=V1, y=V2, shape=Site_field_buff)) + 
  geom_errorbarh(aes(xmax=V1+V1e, xmin=V1-V1e, height=0.003), colour="black") +    
  geom_errorbar(aes(ymax=V2+V2e, ymin=V2-V2e, width=0.003), colour="black") +   
  geom_point(aes(shape = Site_field_buff, alpha = centroid_or_sample, size = centroid_or_sample), stroke=2) +  
  theme_bw() +
  scale_size_manual(values = c(7, 3), name = "Centroid or Sample")+
  scale_alpha_manual(values = c(1, 0.3), name = "Centroid or Sample")+
  #scale_color_manual(values=c("black", "#997700", "#004488"), name="Treatment") +
  scale_shape_manual(values = c(0, 15, 1, 19), name="Site_Buffer") +
  xlab(glue("PCoA 1 ({explainvar1a}%)")) + ylab(glue("PCoA 2 ({explainvar2a}%)")) + 
  geom_segment(data=vec, aes(x=0,xend=V1,y=0,yend=V2), linewidth=1, arrow = arrow(length = unit(0.2, "cm")),colour="black", alpha = 0.6, inherit.aes=F)+ 
 geom_text(data=vec, aes(x=V1, y=V2+0.03), label= vec$labels, size=4, inherit.aes=F)+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
  theme(panel.background = element_blank()) + 
  theme(axis.title = element_text(size=14), axis.text=element_text(size=14), 
          axis.text.x = element_text(size=14),  axis.text.y = element_text(size=14),
          panel.border = element_rect(colour = "black", linewidth=1.25)) +
  theme(axis.ticks.length=unit(0.3,"cm")) + 
  theme(plot.title=element_text(size=14)) +
  theme(legend.text=element_text(size=14), legend.title = element_text(size=14))+
  guides(shape = guide_legend(override.aes = list(size = 4), order = 1), 
           colour = guide_legend(override.aes = list(pch=16, size = 4), order = 2)) +
    ggtitle("")

plot_pcoa

#ggsave("../figures/pcoa.png", plot = plot_pcoa, device="png", path=NULL, scale=1, height = 5, width = 9, limitsize = TRUE, dpi = 600)
```



