---
title: "Town Creek RSC 16S, denitrification, nutrients, and more"
author: "Colin G. Finlay"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    df_print: paged
---
# Setup

```{r setup, include=FALSE}
rm(list=ls())
knitr::opts_chunk$set(echo = FALSE, warning=FALSE)
#use to set working directory 
knitr::opts_knit$set(root.dir="~/GitHub/RSC_Microbes_and_Nutrients/analyses/")
```

# Load packages and functions

```{r load packages and functions, include FALSE }
#Set source R tools
source("../bin/DiversityFunctions.R")
source("../bin/MothurTools.R")

#load required packages
require("vegan")
require("tidyverse")
require("reshape2")
require("ecodist")
require("ggpubr")
require("grid")
require("rstatix")
require("NSM3")
require("ggpmisc")
require("funrar")
require("glue")
require("sf")
require("NSM3")
require("gdm")
require("ARTool")
require("stringr")
```

```{r Bioconductor install and load}
#if (!require("BiocManager", quietly = TRUE))
   # install.packages("BiocManager")
#BiocManager::install(version = "3.22")

require("BiocManager")
```

```{r Phyloseq install and load}
#source("https://raw.githubusercontent.com/joey711/phyloseq/master/inst/scripts/installer.R",
    #   local = TRUE)

#install_phyloseq(branch = "github")

require("phyloseq")

packageVersion('phyloseq')
```

```{r Standard error and confidence intervals}
#Set Std Err and Conf Int
se <- function(x, ...) {
  sd(x, na.rm = TRUE)/sqrt(length(na.omit(x)))
}
ci <- function(x, ...) {
  1.96 * sd(x, na.rm = TRUE)
}
```

# Load files

```{r read in data files}
# Shared file (OTU) from Mothur Februrary 2025 batch run (original shared filed processed in 'CNH2_AgHC_LD_16S.Rmd')
otu_batch <- read.csv("../data/Mothur/OTU/otu_SharedFile_TC_and_TarPam.csv")
```

# QC OTUs per sample

```{r}
# Set row names using mis-formatted column X
rownames(otu_batch) <- otu_batch$X
rownames(otu_batch)

# Remove column X now that row names are set
otu_batch <- otu_batch[,-1]
```

```{r Visualize Town Creek and Tar Pam samples}
# 1. Check dimensions
dim(otu_batch) # 139, 164,672

# 2. Prepare data for plotting
d_otu_4plot <- as.data.frame(otu_batch)
d_otu_4plot$colsum <- rowSums(d_otu_4plot) 
d_otu_4plot$group <- rownames(d_otu_4plot)

# 3. Plot and save OTUs per sample figure
p_TC_OTUs <- ggplot(d_otu_4plot, aes(x = reorder(group, -colsum), y = colsum)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  # Add the count labels
  geom_text(aes(label = colsum), hjust = -0.2, size = 3.5) + 
  coord_flip() +
  theme_minimal() +
  # Expand the y-axis (which is horizontal now) slightly to make room for the labels
  scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
  labs(x = "", y = "", title = "")

p_TC_OTUs

#ggsave("../figures/MothurCheck_Feb26_batch/TC_and_TarPam_OTUs.pdf", plot=p_TC_OTUs, device="pdf", scale=1, width= 200, height= 500, units = "cm", dpi=600, limitsize=F, bg="white")
```
```{r min and max OTU counts}
# 1. Calculate row sums (excluding the first column if it's the ID)
# Assuming SampleID is column 1 and the rest are counts:
total_counts <- rowSums(otu_batch) 

# 2. Find the index of the max and min values
max_index <- which.max(total_counts)
min_index <- which.min(total_counts)

# 3. Extract the Sample IDs using the index
max_sample_id <- rownames(otu_batch)[max_index]
min_sample_id <- rownames(otu_batch)[min_index]

# Print results
print(paste("Largest Sample ID:", max_sample_id, "| Total:", total_counts[max_index])) # GSF1058_TarPam_02 | Total: 400458
print(paste("Smallest Sample ID:", min_sample_id, "| Total:", total_counts[min_index])) # GSF1058_TC_48 | Total: 70
```

```{r Drop all TarPam samples}
# 1. Identify rows that contain "_TarPam_"
# grepl returns TRUE for matches, FALSE for non-matches
is_tarpam <- grepl("_TarPam_", rownames(otu_batch))

# 2. Subset to keep only the rows that do NOT have the pattern
otu_noTarPam <- otu_batch[!is_tarpam, ]

# 3. Sanity Check: Verify the counts
print(paste("Original rows:", nrow(otu_batch))) # 139
print(paste("Rows removed:", sum(is_tarpam))) # 27
print(paste("Remaining rows:", nrow(otu_noTarPam))) # 112
```

```{r Visualize Town Creek samples}
# 1. Check dimensions
dim(otu_noTarPam) # 112, 164,672

# 2. Prepare data for plotting
d2_otu_4plot <- as.data.frame(otu_noTarPam)
d2_otu_4plot$colsum <- rowSums(d2_otu_4plot) 
d2_otu_4plot$group <- rownames(d2_otu_4plot)

# 3. Plot and save OTUs per sample figure
p_TC_OTUs_2 <- ggplot(d2_otu_4plot, aes(x = reorder(group, -colsum), y = colsum)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  # Add the count labels
  geom_text(aes(label = colsum), hjust = -0.2, size = 3.5) + 
  coord_flip() +
  theme_minimal() +
  # Expand the y-axis (which is horizontal now) slightly to make room for the labels
  scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
  labs(x = "", y = "", title = "")

p_TC_OTUs_2

#ggsave("../figures/MothurCheck_Feb26_batch/TC_OTUs.pdf", plot=p_TC_OTUs_2, device="pdf", scale=1, width= 200, height= 400, units = "cm", dpi=600, limitsize=F, bg="white")
```

## Drop Town Creek water samples
Lowest 4 OTU counts are all water (GSF1058_TC_48, GSF1058_TC_50, GSF1058_TC_51, GSF1058_TC_49), and ≤ 425 OTUs. From there, 5th lowest is 7319 OTUs. Plan to drop all water samples before rarefying.

```{r design file for 2015 Town Creek}
# Read in 2015 design file
design_2015 <- read.delim("../data/2015_preRSC/TC.design.txt", row.names = 1)

# Correct Sample IDs and Subset Original Design File
rownames(design_2015) <- gsub("ECU", "GSF1058", rownames(design_2015))
rownames(design_2015) <- gsub("_0", "_", rownames(design_2015))
```


```{r Drop all water samples}
# 1. Identify rows where Sample_Type is water (W)
is_water <- rownames(design_2015)[design_2015$Sample_Type == "W"]

# 2. Subset to keep only the rows that do NOT have the pattern
otu_TC_sed <- otu_noTarPam[!(rownames(otu_noTarPam) %in% is_water), ]

# 3. Sanity Check: Verify the counts
print(paste("Original rows:", nrow(otu_noTarPam))) # 112
print(paste("Rows removed:", sum((rownames(otu_noTarPam) %in% is_water)))) # 34
print(paste("Remaining rows:", nrow(otu_TC_sed))) # 78
```

```{r Visualize Town Creek sediment samples}
# 1. Check dimensions
dim(otu_TC_sed) # 78, 164,672

# 2. Prepare data for plotting
d3_otu_4plot <- as.data.frame(otu_TC_sed)
d3_otu_4plot$colsum <- rowSums(d3_otu_4plot) 
d3_otu_4plot$group <- rownames(d3_otu_4plot)

# 3. Plot and save OTUs per sample figure
p_TC_OTUs_3 <- ggplot(d3_otu_4plot, aes(x = reorder(group, -colsum), y = colsum)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  # Add the count labels
  geom_text(aes(label = colsum), hjust = -0.2, size = 3.5) + 
  coord_flip() +
  theme_minimal() +
  # Expand the y-axis (which is horizontal now) slightly to make room for the labels
  scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
  labs(x = "", y = "", title = "")

p_TC_OTUs_3

#ggsave("../figures/MothurCheck_Feb26_batch/TC_OTUs_SED_ONLY.pdf", plot=p_TC_OTUs_3, device="pdf", scale=1, width= 200, height= 300, units = "cm", dpi=600, limitsize=F, bg="white")
```

## Rarefaction

```{r Gemini rarefaction}
# --- Start here with 'otu_TC_sed' data frame ---

# 1. Find the minimum sample size (library depth) for rarefaction
# We calculate the total counts for each sample (row)
sample_sums <- rowSums(otu_TC_sed)

# We find the smallest sample sum to use as our depth
min_depth <- min(sample_sums)

# It's good practice to print this depth
print(paste("Rarefying all samples to a depth of:", min_depth, "reads"))

# 2. Set a random seed for reproducibility
# Rarefaction involves random subsampling. Setting a seed ensures
# that you get the exact same rarefied data every time you run the code.
set.seed(42) 

# 3. Perform rarefaction
# The rrarefy() function takes the community data (otu2) and
# the subsampling depth (min_depth)
otu_rarefied <- rrarefy(otu_TC_sed, sample = min_depth)

# 4. Check the new row sums to confirm
print("New row sums after rarefaction:")
print(rowSums(otu_rarefied))

# 'otu_rarefied' is the new data frame with rarefied counts
```

Relative abundance using funrar:
- use the package "funrar", and function make_relative() to do relative abundance calculations
  - yields same results as our for loops
  - has publication: https://doi.org/10.1111/ddi.12629 
- install.packages("funrar")
- require("funrar")
- relative_abundance <- make_relative(as.matrix(Numerical))
```{r relative abundance}
otu_rel <- make_relative(as.matrix(otu_rarefied))

# sanity check and TC_RSC count
dim(otu_rel)
# Count row names starting with "TC_RSC"
count <- sum(startsWith(rownames(otu_rel), "TC_RSC"))
count_pre <- sum(startsWith(rownames(otu_rel), "GSF"))
count_pre_2 <- sum(design_2015$Sample_Type %in% c("BD", "BK"))
# Print the result
print(count) # 60
print(count_pre) # 18
count_pre_2 # 18
```

# RSC 2023 denitrification potential

```{r import DEA data}
# Read in denitrification rates from Town Creek DEA, 2023-01-31 sampling:
denit_rates_WINTER <- read.csv("../data/2023_RSC_DEA/DEA_rates/TownCreek_RSC_DEA_20230131.csv")

# Read in denitrification rates from Town Creek DEA, 2023-03-29 sampling:
denit_rates_SPRING <- read.csv("../data/2023_RSC_DEA/DEA_rates/TownCreek_RSC_DEA_20230329.csv")

# Read in denitrification rates from Town Creek DEA, 2023-06-21 sampling:
denit_rates_SUMMER <- read.csv("../data/2023_RSC_DEA/DEA_rates/TownCreek_RSC_DEA_20230621.csv")

# Read in denitrification rates from Town Creek DEA, 2023-12-05 sampling:
denit_rates_FALL <- read.csv("../data/2023_RSC_DEA/DEA_rates/TownCreek_RSC_DEA_20231205.csv")
```

```{r sediment chemical data}
# 2023 RSC sediment N, dry mass corrected:
sediment_N_DM <- read.csv("../data/2023_RSC_DEA/NH4_NOX_per_dryMass.csv")

# 2023 RSC sediment Total Carbon (TC) and Total Nitrogen (TN) from NCSU EAL (2024-10-14):
RSC_sed_TC_TN <- read.csv("../data/2023_RSC_DEA/TownCreekOnly_TC_TN.csv")

# 2023 RSC sediment pH:
sed_pH <- read.csv("../data/2023_RSC_DEA/pH_TC_RSC.csv")

# 2015 sediment and water chem data AND denitrification rates:
WaterSed.chem_2015 <- read.delim("../data/2015_preRSC/TC_WaterSed_ChemDEA.txt")
WaterSed.chem_2015$Field_replicate <- as.factor(WaterSed.chem_2015$Field_replicate)
```

# Wrangle a Meta and OTU data frame

```{r read in meta data}
# 2015 design file has already been read in as 'design_2015'
# Create subset that only contains sediment samples
design_2015_sed <- design_2015 %>%
  filter(Sample_Type %in% c("BD", "BK"))

# Is there a design file for 2023 RSC?
# No, must wrangle

# first, remove acetylene free samples
```

```{r remove NO_ACET}
# Winter:
denit_ACET_WINTER <- dplyr::filter(denit_rates_WINTER, denit_rates_WINTER$Acetylene == "ACET", .preserve = TRUE)

# Spring:
denit_SPRING_ACET <- dplyr::filter(denit_rates_SPRING, denit_rates_SPRING$Acetylene == "ACET", .preserve = TRUE)

# Summer:
denit_SUMMER_ACET <- dplyr::filter(denit_rates_SUMMER, denit_rates_SUMMER$Acetylene == "ACET", .preserve = TRUE)

# Summer:
denit_FALL_ACET <- dplyr::filter(denit_rates_FALL, denit_rates_FALL$Acetylene == "ACET", .preserve = TRUE)
```

```{r write out acetylene only data frames}
# I need to assemble a unified meta df with acetylene only samples to represent the denitrification potential
# Do not include WINTER because no amplicon sequencing data for WINTER

# Spring
#write.csv(denit_SPRING_ACET, file = "../data/2023_RSC_DEA/DEA_rates/ACET_only_DFs/denit_SPRING_ACET.csv")

# Summer
#reduce data frame to two key columns
#denit_SUMMER_ACET_out <- denit_SUMMER_ACET[,c("Sample_ID", "RSC", "Water_column")]
  
#write.csv(denit_SUMMER_ACET_out, file = "../data/2023_RSC_DEA/DEA_rates/ACET_only_DFs/denit_SUMMER_ACET.csv")

# Fall
#reduce data frame to two key columns
#denit_FALL_ACET_out <- denit_FALL_ACET[,c("Sample_ID", "RSC", "Water_column")]
  
#write.csv(denit_FALL_ACET_out, file = "../data/2023_RSC_DEA/DEA_rates/ACET_only_DFs/denit_FALL_ACET.csv")
```

```{r read in and wrangle pre- and post-RSC unified metadata}
# read in data
uni_meta <- read.csv("../data/unified_meta.csv")

# make row names
rownames(uni_meta) <- uni_meta$row_name

# remove row_name column
uni_meta <- uni_meta[,-1]
```

```{r wrangle meta data}
# Cut out "_ACET" from uni_meta
rownames(uni_meta) <- str_remove(rownames(uni_meta), "_ACET$")

# remove barcodes from otu_rel row names
rownames(otu_rel) <- str_remove(rownames(otu_rel), "_bc\\d{3}$")

# Cut out pool name from otu_rel names
# 1. Define the patterns joined by the OR operator "|"
# Notice that "_IS_P4i" comes before "_IS_P4"
strings_to_remove <- "_Dr_P4|_Dr_P5|_IS_P1|_IS_P4i|_IS_P4"

# 2. Remove the variable strings
rownames(otu_rel) <- str_remove(rownames(otu_rel), strings_to_remove)

# In uni_meta, convert submerged and island to BD (stream bed) and BK (stream bank)
uni_meta <- uni_meta %>%
  mutate(Sample_Type = case_match(Sample_Type,
                                  "submerged" ~ "BD",
                                  "island" ~ "BK",
                                  .default = Sample_Type)) # Keeps anything else (like "unknown") as is

# In uni_meta, add a column for pre- vs. post- RSC
uni_meta <- uni_meta %>%
  mutate(PreVsPost = case_match(Location,
                                c("B", "C", "D") ~ "Pre-RSC",
                                c("in-stream", "dry") ~ "Post-RSC",
                                .default = NA_character_))

# In uni_meta, add a column that converts date to categorical season
uni_meta <- uni_meta %>%
  mutate(Season = case_match(Date,
                                c("5/26/15", "3/29/23") ~ "Spring",
                                "6/21/23" ~ "Summer",
                                "12/5/23" ~ "Fall",
                                .default = NA_character_))

# 2. Now cbind the two data frames, which have identical row names and order
meta_otu <- cbind(uni_meta, otu_rel)

# (Optional) Check the dimensions to confirm
dim(meta_otu) # 78 164677
```

```{r set factors}
meta_otu$Sample_Type <- as.factor(meta_otu$Sample_Type)
meta_otu$Location <- as.factor(meta_otu$Location)
meta_otu$PreVsPost <- as.factor(meta_otu$PreVsPost)
meta_otu$Season <- as.factor(meta_otu$Season)
```

# PERMANOVA

```{r PERMANOVA}
metaotu.ad <- adonis2(meta_otu[,-c(1:5)] ~ (PreVsPost + Sample_Type + Season)^2, 
                      method = "bray", 
                      data = meta_otu, 
                      perm = 10000, 
                      by = "terms")

metaotu.ad
```

# RESUME CODING HERE 4/21/2026

```{r PERMDISP}
# 1. Create the distance matrix
dist_matrix <- vegdist(otu_rel, method = "bray")

# 2. Run PERMDISP (betadisper)
# 'groups' should be a vector or factor from your metadata
disper_out <- betadisper(dist_matrix, group = meta_otu$Site)

# 3. Test for significance using a permutation test
perm_test <- permutest(disper_out, permutations = 10000)
print(perm_test) # Insignificant. W/ sig. permanova, indicates true location effect

# 4. Visualize the dispersion
plot(disper_out)
boxplot(disper_out)

# Now repeat with buffer vs. field

# 2. Run PERMDISP (betadisper)
# 'groups' should be a vector or factor from your metadata
disper_out_field_buff <- betadisper(dist_matrix, group = meta_otu$Field_or_Buffer)

# 3. Test for significance using a permutation test
perm_test_field_buff <- permutest(disper_out_field_buff, permutations = 10000)
print(perm_test_field_buff) # Insignificant. W/ sig. permanova, indicates true location effect

# 4. Visualize the dispersion
plot(disper_out_field_buff)
boxplot(disper_out_field_buff)
```

